---
title: "QB Pocket Pressure Analysis (PRSS) — Rolling Window Comparison"
output:
  pdf_document: default
  html_notebook: default
---

## 1. Setup

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(nflfastR)
library(zoo)
library(ggrepel)
```

---

## 2. Load pre-computed RDS data

```{r load-rds, message=FALSE}
# Load pre-computed PRSS results (no need to recompute)
all_summary_k5   <- readRDS("../data/processed/all_summary_k5.rds")
all_summaries    <- readRDS("../data/processed/all_summaries_all_windows.rds")

# Load supporting CSVs (small, fast to read)
plays   <- read_csv("../data/bdb/plays.csv",           show_col_types = FALSE)
players <- read_csv("../data/bdb/players.csv",          show_col_types = FALSE)
pff     <- read_csv("../data/bdb/pffScoutingData.csv",  show_col_types = FALSE)
```

---

## 3. Build PFF pressure labels

```{r pff-pressure}
pff_pressure <- pff %>%
  filter(pff_role == "Pass Rush") %>%
  group_by(gameId, playId) %>%
  summarise(
    actual_pressure = as.integer(any(pff_hit == 1 | pff_hurry == 1 | pff_sack == 1, na.rm = TRUE)),
    n_rushers       = n(),
    .groups = "drop"
  )
```

---

## 4. Merge k=5 summary with plays & PFF

```{r merge-k5}
final_data <- all_summary_k5 %>%
  filter(!is.na(max_neg_change), !is.na(qb)) %>%
  left_join(
    plays %>% select(gameId, playId, possessionTeam, passResult, prePenaltyPlayResult),
    by = c("gameId", "playId")
  ) %>%
  left_join(pff_pressure, by = c("gameId", "playId"))
```

---

## 5. Merge all-windows summary with plays

```{r merge-all-windows}
all_summaries_final <- all_summaries %>%
  filter(!is.na(max_neg_change), !is.na(qb)) %>%
  left_join(
    plays %>% select(gameId, playId, possessionTeam, passResult, prePenaltyPlayResult),
    by = c("gameId", "playId")
  )

# Quick sanity check
all_summaries_final %>%
  group_by(window) %>%
  summarise(
    n         = n(),
    na_count  = sum(is.na(max_neg_change)),
    mean_prss = mean(max_neg_change, na.rm = TRUE),
    sd_prss   = sd(max_neg_change,   na.rm = TRUE)
  )
```

---

## 6. Merge nflfastR play-by-play (EPA, yards)

```{r merge-pbp, message=FALSE}
pbp <- load_pbp(2021) %>%
  mutate(
    gameId = as.numeric(old_game_id),
    playId = as.numeric(play_id)
  ) %>%
  select(gameId, playId, epa, yards_gained, cpoe, air_yards)

all_summaries_final <- all_summaries_final %>%
  left_join(pbp, by = c("gameId", "playId"))

# Confirm join worked
message("Rows with EPA: ", sum(!is.na(all_summaries_final$epa)))
```

---

## 7. Plots

### Plot 1: PRSS Distribution (k=5)

```{r plot-dist-k5}
ggplot(final_data, aes(x = max_neg_change)) +
  geom_histogram(bins = 50, fill = "steelblue", alpha = 0.8) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed") +
  labs(
    title = "Distribution of PRSS (k=5)",
    x     = "PRSS (negative = pocket collapsing)",
    y     = "Count"
  ) +
  theme_minimal()
```

---

### Plot 2: PRSS vs Play Result (k=5)

```{r plot-prss-yards}
ggplot(final_data, aes(x = max_neg_change, y = prePenaltyPlayResult)) +
  geom_point(alpha = 0.2, color = "lightcoral") +
  geom_smooth(method = "lm", color = "black") +
  labs(
    title = "PRSS vs Play Result (yards gained)",
    x     = "PRSS",
    y     = "Yards Gained"
  ) +
  theme_minimal()

cor(final_data$max_neg_change, final_data$prePenaltyPlayResult, use = "complete.obs")
```

---

### Plot 3: PRSS Distribution by Rolling Window Size

```{r plot-density-windows}
ggplot(all_summaries_final, aes(x = max_neg_change, fill = window, color = window)) +
  geom_density(alpha = 0.25) +
  xlim(-20, 10) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  labs(
    title = "PRSS Distribution by Rolling Window Size",
    x     = "PRSS",
    y     = "Density",
    fill  = "Window (k)",
    color = "Window (k)"
  ) +
  theme_minimal(base_size = 13)
```

As the rolling window size increases, the PRSS distribution shifts rightward and becomes wider. k=3 is too reactive to noise; k=10 over-smooths and loses real pressure signals. k=5 or k=7 appears to be the sweet spot.

---

### Plot 4: Correlation with yards by window

```{r corr-yards-window}
all_summaries_final %>%
  group_by(window) %>%
  summarise(
    correlation = cor(max_neg_change, prePenaltyPlayResult, use = "complete.obs"),
    .groups = "drop"
  )
```

---

### Plot 5: PRSS by Pass Result (k=5)

```{r plot-pass-result}
prss_by_result <- all_summaries_final %>%
  filter(window == "5", passResult %in% c("C", "I", "S", "IN")) %>%
  mutate(passResult = factor(passResult,
                             levels = c("S", "IN", "I", "C"),
                             labels = c("Sack", "Interception", "Incomplete", "Complete")))

ggplot(prss_by_result, aes(x = passResult, y = max_neg_change, fill = passResult)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  labs(
    title = "PRSS Distribution by Pass Result (k=5)",
    x     = "Pass Result",
    y     = "PRSS (more negative = more pressure)"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none")

prss_by_result %>%
  group_by(passResult) %>%
  summarise(
    mean_prss   = mean(max_neg_change,   na.rm = TRUE),
    median_prss = median(max_neg_change, na.rm = TRUE),
    n           = n()
  )
```

---

### Plot 6: PRSS vs EPA by window

```{r corr-epa-window}
all_summaries_final %>%
  group_by(window) %>%
  summarise(
    correlation = cor(max_neg_change, epa, use = "complete.obs"),
    n           = n(),
    .groups     = "drop"
  )
```

At the individual play level, PRSS shows very weak correlation with EPA across all window sizes (r ≈ 0.04–0.05). This is expected — pocket geometry is only one of many factors driving play outcomes. Analysis at the QB or team level (aggregating across many plays) is likely more informative.

---

### Plot 7: PRSS Validation vs PFF Pressure Labels (k=5)

```{r plot-validation}
validation_data <- all_summaries_final %>%
  left_join(pff_pressure, by = c("gameId", "playId")) %>%
  filter(!is.na(actual_pressure), window == "5")

ggplot(validation_data,
       aes(x = factor(actual_pressure, labels = c("No Pressure", "Pressure")),
           y = abs(max_neg_change),
           fill = factor(actual_pressure))) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  scale_fill_manual(values = c("steelblue", "tomato")) +
  labs(
    title = "PRSS: Actual Pressure vs No Pressure (k=5)",
    x     = "PFF Pressure Label",
    y     = "PRSS"
  ) +
  theme_minimal(base_size = 13)

t.test(abs(max_neg_change) ~ actual_pressure, data = validation_data)
```

---

### Plot 8: Team PRSS Ranking (k=5)

```{r plot-team-prss}
team_summary <- all_summaries_final %>%
  filter(window == "5", !is.na(possessionTeam)) %>%
  group_by(possessionTeam) %>%
  summarise(
    avg_prss  = mean(abs(max_neg_change), na.rm = TRUE),
    avg_yards = mean(yards_gained,        na.rm = TRUE),
    n_plays   = n(),
    .groups   = "drop"
  ) %>%
  arrange(desc(avg_prss))

ggplot(team_summary, aes(x = reorder(possessionTeam, avg_prss), y = avg_prss)) +
  geom_col(fill = "tomato", alpha = 0.85) +
  coord_flip() +
  labs(
    title = "Avg PRSS by Offensive Team (k=5)",
    x     = "Team",
    y     = "Avg PRSS (higher = more pocket collapse)"
  ) +
  theme_minimal(base_size = 12)
```
